Sizing the bets in a focused portfolio

### 2024-02-23 doc/formal/paper.tex

\title{Sizing the bets in a focused portfolio}
%\title{Capital allocation for a focused portfolio based on fundamental analysis}

@@

The paper attempts to provide a mathematical model and a tool for the focused
investing strategy as advocated by Buffett~\cite{berkshireLetters},
Munger~\cite{almanack}, and others from this investment community.
The approach presented here assumes that the investor's role is to think about
probabilities of different outcomes for a set of businesses. Based on these
assumptions, the tool calculates the optimal allocation of capital for
each of the investment candidates. The model has the option to provide
constraints that ensure: no shorting, no use of leverage, providing a maximum
limit to the risk of permanent loss of capital, and maximum individual
allocation. The software is made available for public use.\\

This paper consist of two parts, one without the mathematics and one with the
mathematics. For a non--mathematical reader, it is best to skip the two sections
related to mathematics and numerics, namely~\autoref{sec:mathematics}
and~\autoref{sec:numerics}. The reader who is interested in mathematics should
still read the preceding sections since they provide context within which the
mathematics is built on. Moreover, for mathematically skilled reader, the
problem formulated and solved here represents a non--linear constrained
optimization problem.

\section{Motivation}
\label{sec:motivation}

Before going on to explain the motivation about this work, a bit of context
might come useful. In silico is an employee--owned engineering consultancy
company that invests its excess cash into publicly traded stocks. The excess
cash comes sporadically, and In silico has been a net buyer over the last four
years. This is a trend that will likely continue. When the excess cash comes in,
the following question arises: How much money to put in which stocks? That
question had been answered by looking at our fundamental analysis for each
company, and performing some vague hand calculations. That of course works well
because of all the other uncertainties in the investment process, but we were
still motivated to:
\begin{enumerate}
    \item Improve the odds of us behaving rationally (i.e. minimize
    psychological misjudgments such as anchoring bias~\cite{kahneman},
    consistency and commitment tendency~\cite{cialdini}, and
    others with their combined effects~\cite{almanack}),
    \item Save time (i.e. avoid doing hand calculations).
\end{enumerate}

Since mathematics is usually good at keeping people rational, and software is
great for saving time, a decision has been made to write a software that answers
the following question: For a set of candidate companies and their current
market capitalization, each having a set of scenarios defined by a probability
and intrinsic value estimate, how much of our capital to invest in each?

@@

The answer to this question has been given by Kelly with his widely known Kelly
formula (sometimes called Kelly criterion)~\cite{kelly}. The Kelly's approach
starts by maximizing long--term growth of capital when one is presented with
an infinite amount of opportunities to bet on. This work extends that idea by
considering:
\begin{enumerate}
    \item Multiple companies (stocks) in parallel,
    \item An arbitrary number of scenarios for each company,
    \item Multiple constraints for modelling our preference towards no shorting,
    no use of leverage, and providing a maximum value for the risk of permanent
    loss of capital,
    \item A fundamentals--based analysis with a very long time horizon. After
    all, a stock is an ownership share of business~\cite{intelligentInvestor},
    in which time--scales are significantly larger than the time--scales of
    day--to--day stock price fluctuations.
\end{enumerate}

The generalization leads to a non--linear system of equations that when solved,
yields a fraction of capital to invest in each of the candidate companies. The
mathematical derivation is presented in \autoref{sec:mathematics}, while in the
next two subsections (\autoref{sec:assumptions} and \autoref{sec:safetyMargin}),
a discussion on assumptions in a non--mathematical way, and the margin of
safety, are presented.

\subsection{Disclaimers and Assumptions}
\label{sec:assumptions}

The underlying philosophy is that one should spend the majority of their time
analyzing investments and thinking about intrinsic values under different
scenarios that might play out, independently of outside thoughts and events.
However, calculating intrinsic value of a company is more of an art than a
science, especially for a high-quality, growing businesses within one's circle
of competence. And according to Charlie Munger, Warren Buffett, Mohnish Pabrai
and the like-minded others from whom the authors got the inspiration for this
work, one should focus precisely on getting such great businesses for a fair
price. That means that one shouldn't take what this approach says at face value,
and one should probably use its guidance infrequently.

During the mathematical derivation presented in \autoref{sec:mathematics}, an
assumption is made that the number of bets is very high (tends to infinity).
This work does not try to justify this assumption in a strong mathematical
form~\footnote{It is on the author's TODO list to try and prove this.}. Here's a
soft, non-mathematical reasoning on why the authors think this is fine: The
assumption is made in order to write the equations in terms of probabilities
instead of the number of outcomes divided by the number of bets. Therefore, as
long as the input probabilities are \textit{conservatively} estimated, the
framework should still be valid. This essentially represents the most important
margin of safety~\cite{intelligentInvestor}.

@@

In addition to the most important margin of safety mentioned above, there are
however a couple of more margins of safety embedded in the current framework.
These are:
\begin{itemize}
    \item No shorting allowed. From a purely mathematical point of view,
    shorting would be allowed. Without detailed math, it is easy to see how a
    company with a negative expected return would result in a short position.
    However, due to the asymmetry of the potential losses compared to gains,
    coupled with the usual time--frame limit that comes with shorting, we
    provided a constraint for being long--only.
    \item No use of leverage allowed. Again, from a purely mathematical point of
    view, use of leverage would sometimes be useful. The thinking in avoiding to
    use leverage is that no--one should be in a hurry to get rich, and should
    avoid risking good night's sleep based on short--term market fluctuations,
    which are fairly hard to predict consistently (unless you work at
    Renaissance Technologies, in which case you can scratch this).
    \item No company without at least one downside scenario is allowed. By
    disallowing inputs without downside, the framework forces you to focus and
    think about what can go wrong, as opposed to dreaming about what can go
    right. This is in--line with thoughts by Prasad~\cite{darwinInvesting} about
    the importance of avoiding the errors of investing in a bad company.
    However, if one is absolutely (100\%) sure that a company does not have a
    downside, the solution is to put all the money in that one company.  It is
    our feeling that the best way to handle such cases is outside of this
    framework. Alternatively, one can always model an unknown downside scenario
    with a small probability (say 5--20\%) and intrinsic value of zero.
\end{itemize}

These assumptions and margins of safety are embedded into the framework as
constraints in order to try and tie the rational mathematics with the real
world.

@@

\noindent If one assumes an infinite number of (re)allocations
$N_a$\footnote{The assumption regarding the infinite number of allocations is
something that the authors are slightly uncomfortable with, but \textit{feel}
it is fine because of the margins of safety embedded into the thinking that goes
into assessing each investment opportunity.}, the following relation holds:
\begin{equation}
\label{eq:7}
    \lim_{N_a \to \infty} \frac{n_i}{N} = p_i
\end{equation}

@@

\indent In this work, we define four constraints that serve as additional
margins of safety in a focused investment approach:
\begin{enumerate}
    \item Long--only constraint ensures that a fraction cannot be negative and
    is added to all candidate companies. Adding this constraint means: "I do not
    allow short positions.":
    \begin{equation}
        \label{eq:11}
            f \ge 0 \rightarrow -f + s = 0
        \end{equation}

    \item Maximum leverage constraint ensures that the leverage is limited up to
    $L$. Note that $L = 0$ implies no leverage. Adding this constraint means: "I
    want to limit the use of leverage (to a minimal amount).":
        \begin{equation}
        \label{eq:12}
            \sum_j^{N_c} f_j \le 1 + L \rightarrow \sum_j^{N_c} f_j - 1 - L + s = 0
        \end{equation}

    \item Maximum individual allocation constraint ensures that a fraction does
    not exceed the specified amount. Adding this constraint means: "I feel
    uncomfortable putting more than $M$ of my capital into a single company.":
        \begin{equation}
        \label{eq:13}
            f \le M \rightarrow f - M + s = 0
        \end{equation}

    \item Maximum allowable permanent capital loss constraint ensures that the
    worst--case outcome does not exceed losing a specified amount of capital
    with a specified probability. Adding this constraint means: "Under a
    worst--case scenario, I am comfortable permanently losing $K$ (e.g. 25\%) of
    my capital with probability $P$ (e.g. 0.1\%).":
        \begin{equation}
        \label{eq:14}
            \sum_j^{N_c} f_j \min(p_{ij} k_{ij}) \ge P \cdot K \rightarrow
            -\sum_j^{N_c} \left(f_j \min(p_{ij} k_{ij}) \right) + P \cdot K + s = 0
        \end{equation}
    \noindent where $\min(p_{ij} k_{ij})$ is the worst--case outcome across all
    scenarios for the $j$-th candidate company. Here, the minimum returns
    $k_{ij}$ and the maximum worst--case return $K$ are both negative by
    convention, indicating a loss, hence the $\ge$ sign.

\indent Note that the maximum allowable permanent capital loss constraint
given by equation~\eqref{eq:14} only works with the long--only constraint
because the fraction $f_j$ is assumed positive. It is also important to note
that here we do not talk about {\it temporary} loss of capital due to short-term
stock market fluctuations, but rather {\it permanent} loss of capital due to the
fundamental business environment of candidate companies. In addition, it is of
course possible for one to lose more than the specified maximum allowable amount
of capital because of the assumptions made with respect to the inputs: After
all, coming up with intrinsic values and probabilities for each scenario is more
of an art than a science.

\end{enumerate}

\subsection{Constrained System: Putting It All Together}
\label{sec:constrainedSystem}

@@

\noindent In order to validate the numerical model, a basic example of five
candidate companies is considered, where each candidate has the same set of
scenarios (probabilities and intrinsic values) and the same market cap.
The inputs that present a 50\% loss and 100\% gain with 50--50 probabilities are
presented in~\autoref{tab:validationCompany}.\\

@@

\begin{table}[!ht]
\caption{Company A with current market cap of 225B USD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss & 0 USD & 5\% \\
Base thesis & 270B USD & 60\% \\
Bull thesis & 420B USD & 35\% \\
\end{tabular}%
\label{tab:companyA}%
\end{table}%

\begin{table}[!ht]
\caption{Company B with current market cap of 450M USD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss& 0 USD & 5\% \\
Bear thesis & 350M USD & 50\% \\
Base thesis & 900M USD & 45\% \\
\end{tabular}%
\label{tab:companyB}%
\end{table}%

\begin{table}[!ht]
\caption{Company C with current market cap of 39M GBP.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss & 0 GBP & 10\% \\
Bear thesis & 34M GBP & 40\% \\
Base thesis & 135M GBP & 50\% \\
\end{tabular}%
\label{tab:companyC}%
\end{table}%

\begin{table}[!ht]
\caption{Company D with current market cap of 751M SGD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Bear thesis & 330M SGD & 30\% \\
Base thesis & 1B SGD & 70\% \\
\end{tabular}%
\label{tab:companyD}%
\end{table}%

\begin{table}[!ht]
\caption{Company E with current market cap of 126B HKD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss & 0 HKD & 5\% \\
Bear thesis & 50B HKD & 10\% \\
Base thesis & 300B HKD & 85\% \\
\end{tabular}%
\label{tab:companyE}%
\end{table}%

\noindent Based on these inputs, with the long--only strategy, without leverage,
and with maximum individual allocation of 30\%, the portfolio allocation that
maximizes the long--term growth--rate of capital is presented
in~\autoref{tab:results}.

@@

\noindent The last item is particularly interesting to the authors. According to
Actuarial Life Tables in~\cite{lifeTables}, the probability of the (currently 34
year--old, the oldest) author dying within the next year is approximately
0.26\%\footnote{This is the only thing that's probably worth measuring in the
basis points in this framework.}. That is two orders of magnitude higher than
the probability of the permanent loss of capital for this portfolio. Considering
that the portfolio has five stocks, that is a very strong argument against
excessive diversification, especially if:
\begin{itemize}
    \item One thinks of stocks as ownership shares of businesses, which implies
    long--term thinking and not being bothered by market fluctuations, 
    \item One embeds a margin of safety in different scenarios for different
    companies by e.g. recognizing that both unknown and known bad things may
    happen.
\end{itemize}

\noindent The observation about excessive diversification is inline with the
thoughts from the Poor Charlie's Almanack~\cite{almanack} and one of the
lectures from Li Lu that the authors frequently watch~\cite{liLuLecture}.

@@

\noindent There are a couple of technical problems that the authors observed:
\begin{enumerate}
    \item It is possible that a given non--linear system for a particular
    combination of constraint statuses (active/inactive) does not converge. This
    may happen if the resulting matrix is singular, or if the Newton--Raphson
    algorithm does not find the solution within a prescribed number of
    iterations. These errors are ignored, which means that it is possible that
    the real best solution is not found due to numerical issues.
    \item The exponential complexity of the model makes it challenging to
    use for more than several candidate companies without using significant
    compute resources. For example, having a 20 candidate companies with all
    constraints would result in around 4 trillion non--linear systems to solve.
    Therefore, the model is not suitable for cases with excessive
    diversification, although there is a possibility to filter some of these
    upfront without attempting to solve them, which may be one of the topics for
    future work.
\end{enumerate}

\indent To conclude, the authors believe that the most challenging aspect of an
investor's work that might use this software is to think hard about the inputs.
The authors see the usefulness of this software mainly in:
\begin{itemize}
    \item Forcing the investors to think consistently in terms of probabilities
    and long--term business outcomes across the range of candidate companies,
    \item Calculating the optimal allocation in a short amount of time, based on
    the investor's inputs.
\end{itemize}

@@

\bibliographystyle{unsrt}
\bibliography{./references}

>>>>

\title{Sizing the bets in a focused portfolio}
%\title{Capital Allocation With Kelly Criterion for a Focused Portfolio Based On Fundamental Analysis}

@@

The paper provides a mathematical model and a tool for the focused
investing strategy as advocated by Buffett~\cite{berkshireLetters},
Munger~\cite{almanack}, and others from this investment community.
The approach presented here assumes that the investor's role is to think about
probabilities of different outcomes for a set of businesses. Based on these
assumptions, the tool calculates the optimal allocation of capital for
each of the investment candidates. The model is based on a generalized Kelly
Criterion with options to provide constraints that ensure: no shorting, limited
use of leverage, providing a maximum limit to the risk of permanent loss of
capital, and maximum individual allocation. The software is applied to an
example portfolio from which certain observations about excessive
diversification are obtained. In addition, the software is made available for
public use.\\

@@

A stock is an ownership share of business~\cite{intelligentInvestor}, in
which time--scales are significantly greater than the time--scales of
day--to--day stock price fluctuations. The work therefore assumes a
fundamentals--based investment analysis with a very long time horizon. There are
generally two parts of a fundamentals--based investment process:
\begin{enumerate}
    \item Finding businesses that are investment candidates,
    \item Deciding how much capital to allocate in each of them.
\end{enumerate}

\noindent This work does not deal with the first part related to finding
relevant investment candidates. It deals only with the second part of the
process, which can be mathematically well formulated, in an attempt to minimize
the effects of the psychological misjudgements such as anchoring
bias~\cite{kahneman}, consistency and commitment tendency~\cite{cialdini}, and
others with their combined effects~\cite{almanack}). Essentially, the question
to answer is: For a set of candidate companies and their current market
capitalization, each having a set of scenarios defined by a probability and
intrinsic value estimate, how much of capital to invest in each?\\

\indent The answer to this question has been given by Kelly~\cite{kelly} with
his widely known Kelly Criterion. Despite the fact that the Kelly Criterion has
been published in 1956 and dealt with information theory and horse racing, its
use is now widespread in the field of finance, with a plethora of recent
research. For example, Pinelis and Ruppret~\cite{pinelisRuppert} noted that the
asset allocation fractions (weights) of the machine--learning powered Maximally
Predictable Portfolio (MPP) are similar to weighting by the Kelly Criterion.
Meister~\cite{meister} used the Kelly Criterion to demonstrate the advantage of
an investor who bets on multiple simultaneous investments compared to an
investor who bets on a single investment. In the limit of infinite number of
simultaneous and similar bets, the first investor has all the capital invested
in a diversified fashion, providing a clear advantage. Byrnes and
Barnett~\cite{byrnesBarnett} went further to generalize the Kelly Criterion with
a continuous probability distribution modeling the price movement of the stock,
while Lototsky and Pollok~\cite{lototskyPollok} rigorously analyzed the
high--frequency limit of the model, concluding that a high--frequency
strategy can lead to a more aggressive betting strategy.  Proskurnikov and
Barmish~\cite{proskurnikovBarmish} investigated the possibility of having a
non--linear control for the Kelly betting, assuming a bounded set for the
input probabilities instead of assuming that they are known in advance.  The
Kelly Criterion was recently even extended to the quantum realm by Meister
and Price~\cite{meisterPrice}, where in addition to the uncertainty related
to flipping a coin, there is an uncertainty in measurement. Perhaps the most
related reference to this work is the paper by Jim\'{e}nez et
al.~\cite{jimenezEtAl} where they extended the Kelly Criterion to both
multivariate and simultaneous bets and applied it to sports betting in the
English Premier League. This work largely follows the approach by
Jim\'{e}nez et al.~\cite{jimenezEtAl}, but applied to a fundamentals based,
focused investing with the following characteristics:
\begin{enumerate}
    \item Multiple companies (stocks) are considered simultaneously,
    \item Each with an arbitrary number of scenarios,
    \item Multiple constraints for modeling authors' preference towards no
    shorting, limited use of leverage, and providing a maximum value for the
    risk of permanent loss of capital,
    \item A fundamentals--based analysis with a very long time horizon, without
    necessarily knowing when the market price will correspond to the
    fundamentals.
\end{enumerate}

The generalization leads to a non--linear system of equations that when solved,
yields a fraction of capital to invest in each of the candidate companies. The
mathematical derivation is presented in \autoref{sec:mathematics}, while in the
next two subsections (\autoref{sec:assumptions} and \autoref{sec:safetyMargin}),
a discussion on assumptions and the margin of safety, are presented.

\subsection{Assumptions}
\label{sec:assumptions}

The underlying philosophy with the fundamentals--based focused investing is that
one should spend the majority of their time analyzing investments and thinking
about intrinsic values under different scenarios that might play out. However,
calculating intrinsic value of a company is a process subject to multiple levels
of uncertainty. Therefore, the reader should always keep in mind that if the
inputs to the mathematical model are unrealistic, the results will be
unrealistic as well.

During the mathematical derivation presented in \autoref{sec:mathematics}, an
assumption is made that the number of bets is very high (tends to infinity).
The assumption is made in order to write the equations in terms of probabilities
instead of the number of outcomes divided by the number of bets. Therefore, as
long as the input probabilities are \textit{conservatively} estimated, the
framework should still be valid. This essentially represents the most important
margin of safety~\cite{intelligentInvestor}.

@@

In addition to the most important margin of safety mentioned above, there are
however a couple of more margins of safety embedded in the current framework.
These are:
\begin{itemize}
    \item No shorting allowed. From a purely mathematical point of view,
    shorting would be allowed. Without detailed mathematics, it is easy to see
    how a company with a negative expected return would result in a short
    position. However, due to the asymmetry of the potential losses compared to
    gains, coupled with the usual time--frame limit that comes with shorting, we
    provided a constraint for being long--only.
    \item Limited use of leverage. Although leverage is especially useful when
    dealing with high--conviction ideas, this constraint is introduced to model
    authors' conservative behavior.
    \item No company without at least one downside scenario is allowed. By
    disallowing inputs without downside, the framework forces us to focus and
    think about what can go wrong, as opposed to focusing on what can go
    right. This is in--line with thoughts by Prasad~\cite{darwinInvesting} about
    the importance of avoiding the errors of investing in a bad company.
    However, if one is absolutely (100\%) sure that a company does not have a
    downside, the solution is to put a significant amount of leveraged capital
    in that one company. There are two options to handle such cases: i) Either
    outside of this framework or ii) Modeling an unknown downside scenario with
    a small probability and intrinsic value of zero.
\end{itemize}

These assumptions and margins of safety are embedded into the framework as
constraints in order to try and tie the mathematical model and the uncertainty
around inputs.

@@

\noindent If one assumes an infinite number of (re)allocations
$N_a$\footnote{The assumption regarding the infinite number of allocations is
somewhat incompatible with a focused investment strategy where a small number of
stocks are considered at any point in time. This incompatibility is addressed by
margins of safety presented in~\autoref{sec:safetyMargin}.}, the following
relation holds:
\begin{equation}
\label{eq:7}
    \lim_{N_a \to \infty} \frac{n_i}{N} = p_i
\end{equation}

@@

\indent In this work, we define four constraints that serve as additional
margins of safety in a focused investment approach:
\begin{enumerate}
    \item Long--only constraint ensures that a fraction cannot be negative and
    is added to all candidate companies. It disallows short positions:
    \begin{equation}
        \label{eq:11}
            f \ge 0 \rightarrow -f + s = 0
        \end{equation}

    \item Maximum leverage constraint ensures that the leverage is limited up to
    $L$. Note that $L = 0$ implies no leverage:
        \begin{equation}
        \label{eq:12}
            \sum_j^{N_c} f_j \le 1 + L \rightarrow \sum_j^{N_c} f_j - 1 - L + s = 0
        \end{equation}

    \item Maximum individual allocation constraint ensures that a fraction does
    not exceed the specified amount. Adding this constraint allows preventing
    excessive concentration in the portfolio:
        \begin{equation}
        \label{eq:13}
            f \le M \rightarrow f - M + s = 0
        \end{equation}

    \item Maximum allowable permanent capital loss constraint ensures that the
    worst--case outcome does not exceed losing a specified amount of capital
    with a specified probability:
        \begin{equation}
        \label{eq:14}
            \sum_j^{N_c} f_j \min(p_{ij} k_{ij}) \ge P \cdot K \rightarrow
            -\sum_j^{N_c} \left(f_j \min(p_{ij} k_{ij}) \right) + P \cdot K + s = 0
        \end{equation}
    \noindent where $\min(p_{ij} k_{ij})$ is the worst--case outcome across all
    scenarios for the $j$-th candidate company. Here, the minimum returns
    $k_{ij}$ and the maximum worst--case return $K$ are both negative by
    convention, indicating a loss, hence the $\ge$ sign.
\end{enumerate}

\indent Note that the maximum allowable permanent capital loss constraint
given by equation~\eqref{eq:14} only works with the long--only constraint
because the fraction $f_j$ is assumed positive. It is also important to note
that here we do not refer to a {\it temporary} loss of capital due to short--term
stock market fluctuations, but rather {\it permanent} loss of capital due to the
fundamental business environment of candidate companies. In addition, it is of
course possible for one to lose more than the specified maximum allowable amount
of capital because of the assumptions made with respect to the inputs.

\subsection{Constrained System}
\label{sec:constrainedSystem}

@@

\noindent In order to validate the numerical model, a basic example of five
candidate companies is considered, where each candidate has the same set of
scenarios (probabilities and intrinsic values) and the same market cap.
The inputs that represent a 50\% loss and 100\% gain with 50--50 probabilities
are presented in~\autoref{tab:validationCompany}.\\

@@

\begin{table}[!b]
\caption{Company A with current market cap of 225B USD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss & 0 USD & 5\% \\
Base thesis & 270B USD & 60\% \\
Bull thesis & 420B USD & 35\% \\
\end{tabular}%
\label{tab:companyA}%
\end{table}%

\begin{table}[!b]
\caption{Company B with current market cap of 450M USD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss& 0 USD & 5\% \\
Bear thesis & 350M USD & 50\% \\
Base thesis & 900M USD & 45\% \\
\end{tabular}%
\label{tab:companyB}%
\end{table}%

\begin{table}[!b]
\caption{Company C with current market cap of 39M GBP.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss & 0 GBP & 10\% \\
Bear thesis & 34M GBP & 40\% \\
Base thesis & 135M GBP & 50\% \\
\end{tabular}%
\label{tab:companyC}%
\end{table}%

\begin{table}[!b]
\caption{Company D with current market cap of 751M SGD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Bear thesis & 330M SGD & 30\% \\
Base thesis & 1B SGD & 70\% \\
\end{tabular}%
\label{tab:companyD}%
\end{table}%

\begin{table}[!b]
\caption{Company E with current market cap of 126B HKD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss & 0 HKD & 5\% \\
Bear thesis & 50B HKD & 10\% \\
Base thesis & 300B HKD & 85\% \\
\end{tabular}%
\label{tab:companyE}%
\end{table}%

\indent Based on these inputs, with the long--only strategy, without leverage,
and with maximum individual allocation of 30\%, the portfolio allocation that
maximizes the long--term growth--rate of capital is presented
in~\autoref{tab:results}.

@@

\noindent The probability of permanent loss of capital of 0.008\% is
particularly interesting. Provided that the inputs are reasonably estimated, and
considering that the portfolio has five stocks, that is a very strong argument
against excessive diversification, especially if:
\begin{itemize}
    \item One thinks of stocks as ownership shares of businesses, which implies
    long--term thinking and not being bothered by market fluctuations, 
    \item One embeds a margin of safety in different scenarios for different
    companies by e.g. recognizing that both unknown and known bad things may
    happen.
\end{itemize}

\noindent The observation about excessive diversification is in--line with the
thoughts from the Poor Charlie's Almanack~\cite{almanack} and one of the
lectures from Li Lu~\cite{liLuLecture}. In addition, it seems that recent
research is also drawing similar conclusions about excessive
diversification~\cite{jamesMenzies,arvanitisEtAl}, although using different
investing frameworks. The study by James and Menzies~\cite{jamesMenzies} is
particularly interesting to the authors since the data suggests that strong
investors should hold concentrated portfolios, which is in--line with practical
observations of extremely successful fundamentals--based investors: Warren
Buffett~\cite{berkshireLetters}, Mohnish Pabrai, Nick Sleep~\cite{nomadLetters},
Pulak Prasad~\cite{darwinInvesting}, and others.

@@

\noindent A couple of problems have been observed by the authors:
\begin{enumerate}
    \item It is possible that a given non--linear system for a particular
    combination of constraint statuses (active/inactive) does not converge. This
    may happen if the resulting matrix is singular, or if the Newton--Raphson
    algorithm does not find the solution within a prescribed number of
    iterations. These errors are ignored, which means that it is possible that
    the optimal solution is not found due to numerical issues.
    \item The exponential complexity of the model makes it challenging to
    use for more than several candidate companies without using significant
    compute resources. For example, having a 20 candidate companies with all
    constraints would result in around 4 trillion non--linear systems to solve.
    Therefore, the model is not suitable for cases with excessive
    diversification, although there is a possibility to filter some of these
    upfront without attempting to solve them, which may be one of the topics for
    future work.
\end{enumerate}

\indent To conclude, we believe that the most challenging aspect of an
investor's work that might use this software is to appropriately and
conservatively model the inputs. The authors see the usefulness of this software
mainly in:
\begin{itemize}
    \item Forcing the investors to think consistently in terms of probabilities
    and long--term business outcomes across the range of candidate companies,
    \item Avoiding psychological biases by having a tool that calculates the
    optimal allocation of capital based on a strict mathematical formalism.
\end{itemize}

@@

\bibliographystyle{biolett}
\bibliography{./../references}
